---
title: "Reading tabular data from HDF5"
author: "Marc Paterno"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading tabular data from HDF5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(h5tables)
root <- rprojroot::is_r_package

```

The `h5tables` packages provides functions to read `tibbles` from *tabular HDF5 files*.

A tabular HDF5 file is one that contains, at the top-level, on Groups,
and in which each Group contains only Datasets.
Within each Group, each Dataset is of the same length.
Thus each Group represents a *table*, and each Dataset a *column* in a table.

## Discovering what is in a file

To list all the tables in a tabular HDF5 file, use `h5tables::table_names`:

```{r}
names <-table_names(root$find_file("inst/extdata/cms_two_events.h5"))
names
```


## Reading tables

Tables can be read *whole* (all columns, all rows),
or by selecting the names of columns (some columns, all rows).